Probabilistic Network Latency Verification

### 2023-02-06 paper/2_overview.tex
[Paper] Completed section 2

\subsection{Timed Automata}
Timed Automata is an extension to the classic Finite State Machine (FSM) in automata theory that 
adds a global clock to be used as constraints. 
The runtime will keep a set of global clock that monotonously increases and can be individually 
reset.
We could construct the state machine that constraints each transition with a time limit. 
By having this additional time constraints, we could verify properties in a similar way 
as traditional model checking methods (e.g. using CTL formulas).

@@

However, they did it by collecting the statistical property of multiple simulation runs, which 
has the same drawback as the simulation-based method. 
It is also limited in its choice of distribution, only allowing user to use uniform and 
exponential distribution.

@@

One major downside of this framework, however, is the assumptions about what kind of probability 
distributions being used. 
There is a closed-form solution for a network where the queues are identified as a poisson process 
(i.e. Jackson network). 
Still, a queuing network with a more diverse queuing behavior does not have as strong theoretical 
answer in this framework.

\subsection{Convolution}

@@

We divide the problem of latency verification into two parts: 
verifying that two nodes are reachable (\textbf{functional verification}) and only if 
the functional property is fulfilled, we would verify whether the latency between 
two nodes fulfilled some temporal condition (\textbf{temporal verification}).

>>>>

\subsection{Timed Automata}
Timed Automata is an extension to the classic Finite State Machine (FSM) in automata theory that 
adds a global clock to be used as transition constraints. 
The runtime will keep a set of global clock that monotonously increases and can be individually 
reset.
Using these clocks, we could construct the state machine that constraints each transition with 
a time limit. 
With this additional time constraints, we could verify properties in a similar way 
as traditional model checking methods (e.g. using CTL formulas).

@@

However, they did it by collecting the statistical property of multiple simulation runs, which 
has the same drawback as the simulation-based method. 
It is also limited in its choice of probability distribution, only allowing user to use uniform and 
exponential distribution.

@@

One major downside of this framework, however, is the limitation of assumed probability 
distributions that is being used in each queue and the arrival process. 
Where one could derive a closed-form solution for a network where the queues and / or the arrival process 
are identified as a poisson process (i.e. Jackson network, BCMP network \cite{bcmp}), 
the same thing cannot be said for a queuing network that posits a more diverse queuing behavior.
A framework to describe a queuing network with an arbitrary queuing behavior and arrival process in this 
theory is yet to be discovered.

\subsection{Latency as Random Variable}
% Verification framework that is
% - Expressive (congestion control, load-balancing scheme, traffic generation -> probabilistic)
% - Performant (verification time that makes sense, proper-level of abstraction)
% - Practical (reasonably accurate while requiring minimal amount of information and assumption)

Considering these available approaches and their respective strength and limitation, we devise a 
simple yet powerful framework to verify the latency property in a given network.

Our main idea is to model the latency of a given network component as an independent random variable.
By modeling the latency this way, we could then operate on those random variable to model the latency 
of a given path, and eventually the latency of a given src-dst pair.
Using the resulting random variable, we could then determine the temporal property in question by 
analyzing the statistical property of the resulting probability distribution.

\textbf{This framework is expressive}.
The packet interarrival time that is represented by the latency distribution in each router encodes the 
latency information of the various traffic that is going through the network.
At the same time, it also encodes the effect of congestion and flow control scheme that affects 
those traffic.

\textbf{This framework is performant}.
Operating on the high-level of abstraction of the latency distribution in each network components allows 
this framework to perform the verification task in a matter of minutes?
We demonstrate our evaluation result on Section \ref{eval}.

\textbf{This framework is practical}.
Queue length is a common metric that is used to measure the resource utilization of a network 
\cite{dctcp} \cite{swift}.
Using this metric, one could easily construct a latency distribution by combining it with the 
line rate bandwidth.

@@

We divide the problem of latency verification in this framework into two parts: 
verifying that two nodes are reachable (\textbf{functional verification}) and only if 
the functional property is fulfilled, we would verify whether the latency between 
two nodes fulfilled some temporal condition (\textbf{temporal verification}).
In Section \ref{sec:functional}, we briefly describe the functional verification scheme that 
our framework is built upon. 
Next, in Section \ref{sec:temp}, we describe our novel temporal verification framework 
in more detail.
Next, in Section \ref{sec:opt}, we describe the optimization techniques that arise from 
these two steps.
We then evaluate our implementation in Section \ref{eval}.
Finally, we will touch on some related works on Section \ref{sec:rel}, limitations and future 
works on \ref{sec:fut}, and conclude our work with Section \ref{sec:conc}.

### 2023-02-06 paper/3_functional.tex
[Paper] Completed section 2

\section{Functional Verification}
For functional property verification, NetDice \cite{netdice} 
has laid the way for verifying reachability between two nodes under failure in a 
probabilistic and efficient manner.
To contextualize our modification, we will briefly describe some parts of 
NetDice's network encoding and exploration algorithm as a prelude to our alteration.
For a more complete complete description of NetDice, please refer to the original 
paper.

>>>>

\section{Functional Verification} \label{sec:functional}
For functional property verification, NetDice \cite{netdice} 
has laid the way for verifying reachability between two nodes under failure in a 
probabilistic and efficient manner.
To contextualize our modification, we will briefly describe some parts of 
NetDice's network encoding and exploration algorithm as a prelude to our alteration.
For a more complete complete description of NetDice, please refer to the original 
paper.

### 2023-02-06 paper/4_temporal.tex
[Paper] Changed Intro
[Paper] Completed section 2

\section{Temporal Verification}
In the temporal verification stage, we will use the path information that we got from 
functional verification stage and augment them with latency information to 
produce the final temporal property probability.

@@

\section{Optimization}
\subsection{Equivalence Class Consolidation}
To minimize the amount of temporal verification overhead, we want to find some symmetry in the 
existing exploration algorithm so that we could reduce the exploration space even further.

@@

% \textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
% \cite{steffen2020probabilistic} have done. 
% In addition, we also collect additional information from each state (e.g. convergent 
% paths) to be used in the temporal verification step.

>>>>

\section{Temporal Verification} \label{sec:temp}
In the temporal verification stage, we will use the path information that we got from 
functional verification stage and augment them with latency information to 
produce the final temporal property probability.

@@

\section{Optimization} \label{sec:opt}
\subsection{Equivalence Class Consolidation}
To minimize the amount of temporal verification overhead, we want to find some symmetry in the 
existing exploration algorithm so that we could reduce the exploration space even further.

@@

% \textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
% \cite{netdice} have done. 
% In addition, we also collect additional information from each state (e.g. convergent 
% paths) to be used in the temporal verification step.
